\documentclass[12pt]{article}
\usepackage{amsmath}
\newcommand{\bra}[1]{\langle#1|}
\newcommand{\ket}[1]{|#1\rangle}
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\beqa}{\begin{eqnarray}}
\newcommand{\eeqa}{\end{eqnarray}}
\begin{document}

\title{Calculating the Gradient of a Cost Function\\
 for a Parametric Quantum Circuit in\\
 FIVE EASY PIECES}


\author{Robert R. Tucci\\
        P.O. Box 226\\
        Bedford,  MA   01730\\
        tucci@ar-tiste.com\\
        www.artiste-qb.net}


\date{\today}
\maketitle
\vskip2cm

Hybrid Quantum Classical (HQC) computation
as being pursued by Rigetti Inc. involves
minimizing the mean value of a Hermitian operator,
wherein that mean value is calculated empirically from
the data yielded by Rigetti's qc.
One very popular method of minimizing
a function is using back propagation (BP).
The software PennyLane by Xanadu Inc. and the software
Qubiter that I manage for Artiste-qb.net Inc.
can already do BP
to a limited extent
on the Rigetti virtual and real qc's.
The goal of this brief article is to
discuss some ideas
that might allow us
to exploit BP
to its full potential in HQC in the future.

Suppose that you want to  minimize a cost function
$C(x)$ where $x$ has components $x_k$. BP is a way of calculating
$\nabla C(x)$ which one then uses to update
 the value of $x$ iteratively
using what is called gradient descent:

\beq
x^{new}_k = x_k - \eta\frac{\partial C(x)}{\partial x_k}
\;,
\eeq
where $\eta>0$.
The idea behind gradient descent is
 that the increment in cost is
$dC= (x^{new}_k-x_k)(dx_k)(-1/\eta)$,
which we expect to be negative
since $(x^{new}_k-x_k)(dx_k)\approx (dx_k)^2>0$.
If $dC<0$ each time, then we expect the cost will
 move towards a minimum.
At this point, the question arises,
what $C(x)$ should one use for HQC and how
does one calculate it
and its gradient?

Qubiter currently minimizes $C(x)=C_{exact}(x)$,
where $C_{exact}(x)$ is the cost function
calculated exactly (theoretically)
from the wavefunction calculated
by the Qubiter simulator. This
is an interesting case because,
to calculate $\nabla C_{exact}$,
one can do BP distributively, using GPU and TPU,
via softwares like TensorFlow and PyTorch.
It is known that the computational
complexity of back propagation (BP) and forward
propagation (FP) are about the same.
Since doing a classical simulation
of a quantum circuit (in other words, doing FP)
blows up exponentially with the number of qubits,
the same will be true if we attempt to do BP.
So why do either?
The motivations for doing BP
to calculate $\nabla C_{exact}$
on a quantum simulator are the same as those
for doing FP on a quantum simulator.
One does it to generate new ideas and
to test various things
on a smallish number of qubits.

Ultimately, the
HQC people will want to minimize $C(x)=C_{emp}(x)$,
instead of $C(x)=C_{exact}(x)$,
where $C_{emp}(x)$ is the cost function calculated
empirically, from the data yielded by the qc
hardware. Presumably, qc
hardware
can calculate $C_{emp}(x)$
and its gradient
much faster
that classical hardware
can calculate $C_{exact}(x)$
and it gradient.
Calculating the gradient
of $C_{emp}(x)$
has many potential
benefits,
but it is not
obvious
what is the best way of doing this.
Finding a good way will require
some novel and
careful thinking. If one does
something like BP,
this will require calculating the derivatives
$\frac{\partial}{\partial x_k}$ of
each gate that depends on the parameters $x$
of the quantum circuit.
$C_{emp}$ is a statistical quantity
compiled from many ``shots" (samples), so
it fluctuates, and a naive calculation of
the derivatives of a gate
by a finite difference method,
with no other type of averaging, is bound to fail.
It's difficult to calculate meaningfully
the difference of two values that
are very close and fluctuating.

Next, I will show how to calculate
the derivative of
a quantum gate with respect
to one of its parameters,
by calculating 5 separate mean values,
and obtaining the derivative
as a linear combination of those 5 mean values.
The authors of PennyLane
have come up with a similar
scheme, but their method
is different to mine.
To tackle a
general U(2)
transformation,
they first decompose
it
into an Euler
product of 3 rotations along
the standard X,Y or Z axes,
and then they take the derivatives
of each of those 3 rotations.
In this paper,
I
give a
method that can
handle an arbitrary
U(2)
transformation,
without
having to
do the Euler
decomposition first.
Nevertheless, check their method out!
You might like it more than the
method I propose here.

Lucky for us,
the parameters of a quantum gate
almost always appear inside
a 2-dim unitary matrix (an element
of the group U(2)).
So, from here on, we will only concern ourselves
with calculating the derivatives
of a U(2) matrix.


Let's start easy, with a rotation about
a standard axis, $X, Y, Z$,
instead of a general U(2) matrix.
If we let
 $\sigma_k$ for $k=1,2,3$ denote the Pauli matrices,
 then a rotation about the $Z$ axis
 is

\beq
U(\theta_3) = e^{i\sigma_3\theta_3} = C + i\sigma_3 S
\;,
\eeq
where
$\theta_3$
is some real number and
we abbreviate $S = \sin\theta_3, C = \cos \theta_3$.
Then

\beq
\frac{dU}{dt} = \dot{\theta}_3(-S + i\sigma_3 C)
\;.
\eeq
Hence
\beq
\frac{dU}{d\theta_3}=
-S + i\sigma_3 C=e^{i(\frac{\pi}{2}+\theta_3)\sigma_3}
=U(\frac{\pi}{2}+\theta_3)
\;.
\eeq
Thus, for a rotation along a standard axis,
one can evaluate the derivative of a gate
simply
by replacing that gate by that gate
with its angle advanced by $\frac{\pi}{2}$.
No need to take finite differences.
This begs the question,
can we calculate the gradient
of a general U(2) matrix,
in an exact, closed form
that is just as
convenient? Yes we can, as I will show next.

Now let us consider the most general U(2).
We will parameterize it as

\beq
U= e^{i(\theta_0 + \theta_1\sigma_X
+ \theta_2\sigma_Y
+ \theta_3\sigma_Z)}
\;,
\label{u2_pametrization}
\eeq
where $\theta_k$ for $k = 0, 1, 2, 3$ are real numbers.
Derivatives with respect to $\theta_0$
are trivial so we will set $\theta_0=0$ henceforth.
Expressing things using the Einstein summation convention,

\beq U = e^{i\sigma_k\theta_k} = C +
i\sigma_k \frac{\theta_k}{\theta} S
\;,
\eeq
where we are abbreviating

\beq
\theta = \sqrt{\theta_k\theta_k},
 S = \sin\theta, C = \cos \theta
 \;.
 \eeq
 Then, it's easy to show that

 \beq
 \frac{dU}{dt}=-S \frac{\theta_k}{\theta}
 \dot{\theta_k}+ i\sigma_k\dot{\theta_r}
 \left[\frac{\theta_k\theta_r}{\theta^2} C+
  \frac{S}{\theta}(-\frac{\theta_k\theta_r}{\theta^2}
  + \delta_{k, r})\right]
  \;.
  \label{gen_der}
  \eeq
 Eq.(\ref{gen_der}) has
 been checked numerically
 by Qubiter's code. and is already
coded into Qubiter's implementation of Autograd.


In the rest of this article,
we will try to recast the right hand side of
 Eq.(\ref{gen_der}) into a form that
 is more convenient for empirical
 calculation from qc data.
Before embarking on this task,
let us introduce some notation.
As physicists are fond of doing, we will
represent a unit vector by
a letter with a caret above it:
$\hat{a} = \frac{\vec{a}}{|\vec{a}|}$.
Also, for any 3-dim vector $\vec{a}$, let

\beq
\sigma_{\vec{a}}=\vec{a}\cdot\vec{\sigma}
\;.
\label{paulions}
\eeq
Eq.(\ref{paulions}) is a natural generalization
of the Pauli matrix notation. If $\hat{e}_A$
is the unit vector in direction $A$ for $A=X,Y,Z$,
then $\hat{e}_A\cdot\vec{\sigma} = \sigma_A$ for
$A=X,Y,Z$. Expressed in
the notation of Eq.(\ref{paulions}),
two familiar Pauli matrix identities
are

\beq
\sigma_{\vec{a}}\sigma_{\vec{b}}=\vec{a}\cdot\vec{b}
+ i\sigma_{\vec{a}\times\vec{b}}
\;,
\eeq
where $\vec{a}$ and $\vec{b}$
are any two 3-dim vectors,
and

\beq
e^{i\theta\sigma_{\hat{n}}} = \cos\theta +
i\sigma_{\hat{n}}\sin\theta
\;,
\label{cis}
\eeq
where $\theta$ is a real number and $\hat{n}$
is a 3-dim unit vector. Eq.(\ref{cis})
can be proven by Taylor expanding,
and using $\sigma_{\hat{n}}^2=1$.

We will also use the following notation for the projectors
$P_0$ and $P_1$
along the direction of $\ket{0}$
and $\ket{1}$, respectively,
in a 1-qubit space.

\beq
\begin{array}{l}n=P_1=|1\rangle\langle 1|,\\
\bar{n}=1-n=P_0=|0\rangle\langle 0|
\end{array}
\;.
\eeq
$n$ is often called the number operator.
Whenever we say $\Omega(\alpha)$
for a 1-qubit operator $\Omega$,
we mean $\Omega$ applied to qubit $\alpha$.


As usual,
let $U\in U(2)$.
Our next goal is to express , $\dot{U}$,
the derivative
of $U$  with respect to a
parameter $t$, as a linear
combination
of unitary operators $U_k$,
where the real numbers $x_k$ sum to one:

\begin{subequations}
\label{ideal_udot_form}
\beq
\dot{U}=\sum_k x_k U_k
\;,
\eeq


\beq
\sum_k x_k = 1,\;\;U_kU_k^\dagger=1 \;\;\;\forall k
\;.
\eeq
\end{subequations}
It's important to note that
we don't require $x_k>0$ for all $k$,
so the $x_k$ are not probabilities.

Why do we want to express $\dot{U}$
in the form of Eqs.(\ref{ideal_udot_form})?
Because in a qc, $\dot{U}$
will often be subject to 1 or more controls.
Controls
are easy to deal with
if $\dot{U}$
can be expressed as
in Eqs.(\ref{ideal_udot_form}),
because then

\beq\dot{U}(0)^{n(1)n(2)}=
\left[\sum_k x_k U_k\right]^{n(1)n(2)}
=
\sum_k x_k U_k(0)^{n(1)n(2)}
\;.
\label{easy_controls}
\eeq
The fact that Eq.(\ref{easy_controls})
holds can be easily verified
by considering the two
cases $n(1)n(2)=0,1$ separately.
If LHS=left hand side,
RHS=right hand side, refer to the two sides of
Eq.(\ref{easy_controls}):


\beq
\begin{array}{lll}
n(1)n(2)=0: &LHS=1,& RHS=\sum_k x_k=1\\
n(1)n(2)=1:&LHS= \dot{U}(0),&RHS= \sum_k x_k U_k(0)
\end{array}
\;.
\eeq

Eq.(\ref{gen_der}) for the general form of
$\dot{U}$ when parameterized
as Eq.({\ref{u2_pametrization}) (with $\theta_0=0$)
is fairly opaque.
To clarify
Eq.(\ref{gen_der}),
we start by specializing it to $t=\theta_1$.
The cases $t=\theta_2, \theta_3$
can be obtained from the case
$t=\theta_1$ simply by replacing
1 subscripts in our final result by 2 or 3.
So, setting $t=\theta_1$ in Eq.(\ref{gen_der}),
we immediately get:

\beq
\frac{\partial U}{\partial \theta_1} =
\left\{
\begin{array}{l}
-\frac{\theta_1 S}{\theta^2}
\left[i\sigma_{\hat{\theta}}\right]
\\
+
\frac{\theta_1}{\theta}
\left[-S + i\sigma_{\hat{\theta}}C\right]
\\
+
\frac{S}{\theta}
\left[i\sigma_1\right]
\end{array}
\right.
\;.
\eeq
If we define

\beq
p_1=\frac{\theta_1}{\theta},\;\;p_S = \frac{S}{\theta}
\;,
\eeq
then

\beq
\frac{\partial U}{\partial \theta_1} =
\left\{
\begin{array}{l}
-p_1p_S
\left[e^{i\frac{\pi}{2}\sigma_{\hat{\theta}}}\right]
\\
+
p_1
\left[e^{i(\frac{\pi}{2}+\theta)\sigma_{\hat{\theta}}}\right]
\\
+
p_S
\left[e^{i\frac{\pi}{2}\sigma_1}\right]
\end{array}
\right.
\;,
\eeq
which is equivalent to

\beq
\boxed{
\frac{\partial U}{\partial \theta_1} =
\left\{
\begin{array}{l}
\frac{1}{2}(1-p_1)(1-p_S)[1]
\\
+
\frac{1}{2}(1-p_1)(1-p_S)[-1]
\\
-p_1p_S
\left[e^{i\frac{\pi}{2}\sigma_{\hat{\theta}}}\right]
\\
+
p_1
\left[e^{i(\frac{\pi}{2}+\theta)\sigma_{\hat{\theta}}}\right]
\\
+
p_S
\left[e^{i\frac{\pi}{2}\sigma_1}\right]
\end{array}
\right.
\;.}
\label{fin_decomp_of_udot}
\eeq
But note that

\beq
p_1 + p_S -  p_1p_S + (1-p_1)(1-p_S) = 1
\;.
\eeq
If we let $U_k$
equal the
unitary matrices
inside the square brakets in
the RHS of Eq.(\ref{fin_decomp_of_udot}),
then it is clear that
Eq.(\ref{fin_decomp_of_udot})
satisfies

\begin{subequations}
\beq
\frac{\partial U}{\partial \theta_1}=
\sum_{k=1}^5 x_k U_k
\;,
\eeq
and

\beq
\sum_{k=1}^5x_k=1,\;\;\;
U_kU_k^\dagger=1\;\;\;\forall k
\;.
\eeq
\end{subequations}
Just what we wanted!
$\frac{\partial U}{\partial \theta_1}$
in five easy pieces. And replace 1 by 2 or 3
in Eq.(\ref{fin_decomp_of_udot}) to get
partials of $U$ with respect to $\theta_2$
and $\theta_3$.

\section*{ADDED LATER}
The above essay was written in a colloquial style
with the aim of generating interest among
 my peers in this subject.
A week later, I am adding this more technical
appendix to
fill in some of the gaps that
were
left behind in the above colloquial treatment.
What I do in this appendix is to show how to
express the
gradient of the cost function as a sum of
mean values that are
readily evaluated empirically on a real qc.

Suppose $a, A$ are integers such that $a\leq A$,
and $\Omega_j$ are operators acting on $N_b$ qubits.
Define
\beq
\Omega_{[a,A]} = \Omega_a \Omega_{a+1}\ldots \Omega_{A}
\;
\eeq
and

\beq
\Omega_{[A,a]} = \Omega_A\ldots \Omega_{a+1} \Omega_{a}
\;.
\eeq
Note that

\beq
(\Omega_{[a, A]})^\dagger = \Omega^\dagger_{[A,a]}
\;.
\eeq

The cost function that we minimize
in Hybrid Quantum Classical computing
 can be expressed analytically as

\beq C = \bra{\psi_0}(\Omega_{[T-1,0]})^\dagger H
\Omega_{[T-1,0]}\ket{\psi_0}
\;,
\eeq
where the $\Omega_j$ are unitary operators
and $H$ is a Hermitian operator
acting on $N_b$ qubits. Now let us
focus on
taking the derivative of $\Omega_\tau$, the gate for a particular time
$\tau \in \{0,1, \ldots T-1\}$. Define

\beq
\ket{\psi_\tau} = \Omega_{[\tau-1, 0]}\ket{\psi_0}
\;
\eeq
and

\beq
H_\tau = (\Omega_{[T-1,\tau]})^\dagger H \Omega_{[T-1,\tau]}
\;.
\eeq
Henceforth,
we will
use the angled brackets to
 denote an average with respect to
state $\ket{\psi_\tau}$:
\beq
\bra{\psi_\tau}\cdot\ket{\psi_\tau}
=
\langle \cdot\rangle
\;.
\eeq
Using the above notation, the
cost function and its derivative
with respect to a parameter $\theta_{\tau 1}$
that lives inside the operator $\Omega_{\tau 1}$,
can be expressed as

\beq
C = \langle H_\tau\rangle
\;,
\eeq
and


\beq
\frac{\partial C}{\partial\theta_{\tau 1}}
=
\langle H_\tau
\Omega_\tau^\dagger
\frac{\partial \Omega_\tau}{\partial\theta_{\tau 1}}
\rangle
+ h.c.
\;
\label{derv_cost1}
\eeq
h.c. denotes the hermitian conjugate
of the preceding twin expression.

Let $U_{\tau 1}(0)$
be an element of $SU(2)$ acting on qubit 0.
$U_{\tau 1}(0)$ can be parameterized as


\beq
U_{\tau 1}(0)=
e^{i[\theta_{\tau 1}\sigma_X(0)
+
\theta_{\tau 2}\sigma_Y(0)
+
\theta_{\tau 3}\sigma_Z(0)]}
\;,
\eeq
where the $\theta_{\tau d}$ for $d=1,2,3$
are real numbers,
and $\sigma_X(0), \sigma_Y(0), \sigma_Z(0)$
are
the Pauli matrices acting on qubit 0.
For definiteness and as a good illustration,
we will henceforth assume that the
unitary operator $\Omega_\tau$
has the special form
of an SU(2) gate with two controls:

\beq
\Omega_{\tau}=
U_{\tau 1}(0)^{n(1)n(2)}
\;,
\label{special_omega_tau}
\eeq
where, as usual, $n=\ket{0}\bra{0}$
is the number operator, and $n(1), n(2)$
are number operators acting on qubits 1 and 2,
respectively.
As was shown in the main part of this essay,
one can express the partial derivative of
$U_{\tau 1}(0)$ with respect
 to its parameter $\theta_{\tau 1}$,
as a linear combination


\beq
\frac{\partial U_{\tau 1}(0)}{\partial\theta_{\tau 1}}
=
\sum_k \lambda_{\tau 1,k} V_{\tau 1,k}(0)
\;,
\label{lin_comb}
\eeq
where the
coefficients $\lambda_{\tau 1,k}$
are real numbers and the
$V_{\tau 1,k}(0)$ are elements of $SU(2)$.
Here
$k=1,2,3$ (we exclude the two terms that sum to zero
because there is no need for them in this situation).
From Eqs.(\ref{special_omega_tau})
and (\ref{lin_comb}), it follows that

\beq
\Omega^\dagger_\tau
\frac{\partial \Omega_\tau}{\partial\theta_{\tau 1}}
=
\sum_k \lambda_{\tau 1,k}\left[
n(1)n(2)U^\dagger_{\tau 1,k}(0)V_{\tau 1,k}(0)
\right]
\;,
\eeq
which, when substituted into Eq.(\ref{derv_cost1}),
yields:

\beq
\frac{\partial C}{\partial\theta_{\tau 1}}
=
\sum_k \lambda_{\tau 1,k}
\left\langle
H_\tau n(1)n(2)U^\dagger_{\tau 1,k}(0)V_{\tau 1,k}(0)
+h.c.\right\rangle
\;.
\eeq

Note that
$U^\dagger_{\tau 1,k}(0)V_{\tau 1,k}(0)$ is
a product of SU(2) matrices, so
it is itself an SU(2)
matrix acting on qubit 0.
Hence, it can be expressed as

\beq
U^\dagger_{\tau 1,k}(0)V_{\tau 1,k}(0)
= c_{\tau 1,k} +i \sigma_{\hat{\alpha}_{\tau 1,k}}(0)
s_{\tau 1,k}
\;,
\eeq
where we abbreviate

\beq
c_{\tau 1,k} = \cos(\alpha_{\tau 1,k}), \;\;
s_{\tau 1,k} = \sin(\alpha_{\tau 1,k})
\;.
\eeq
$\alpha_{\tau 1,k}$ is a real number
and $\hat{\alpha}_{\tau 1,k}$ is a real
valued unit vector.

Note also that, since

\beq
n(1)  = \frac{1-\sigma_Z(1)}{2}, \;\;
n(2)  = \frac{1-\sigma_Z(2)}{2}
\;,
\eeq
one has

\beqa
n(1)n(2) &=&
\frac{1}{4}[1 - \sigma_Z(1) - \sigma_Z(2)
+ \sigma_Z(1)\sigma_Z(2)]\\&=&
\frac{1}{4}\sum_{a=0}^1\sum_{b=0}^1 (-1)^{a+b}
 \sigma^a_Z(1)\sigma^b_Z(2)
\;.
\label{n1n2_sum}
\eeqa

At this point, it clear that
in order to evaluate
$\frac{\partial C}{\partial\theta_{\tau 1}}$
empirically, one needs to
evaluate empirically two
types of mean values
that we will refer to as types A and B:

\beq
A_\tau = \langle H_\tau n(1)n(2) + h.c.\rangle
\;,
\eeq

\beq
B_{\tau,k} = \langle H_\tau \sigma^a_Z(1)\sigma_Z^b(2)
i \sigma_{\hat{\alpha}_{\tau 1,k}}(0) + h.c.\rangle
\;.
\eeq
Note that if we use Eq.(\ref{n1n2_sum})
in type A (as we did for type B),
then the two terms on the RHS of type A
are each a product of Pauli matrices,
whereas in type B, the two terms on the RHS
are each a product of Pauli matrices \underline{times $i$}.
As we shall see, that extra $i$ in type B
has giant repercussions, making very different
the methods that can be used to evaluate  type A and B
mean values. As we shall see, evaluating
type A mean values
requires some post-selection, whereas
evaluating type B ones doesn't.
Type A mean values don't arise if
gate $\Omega_\tau$ has no controls.
So far, the PennyLane software only
considers evaluating the gradient of uncontrolled
gates, so they never have to evaluate
type A mean values.

The above definitions for type A and B
mean values are not in a form that is
readily evaluated empirically on a qc.
The rest of this note will be devoted to
recasting them into a new form that is.


Define
\beq
\Sigma(0,1,2) = \sigma^a_Z(1)\sigma_Z^b(2)
 \sigma_{\hat{\alpha}_{\tau 1,k}}(0)
\;.
\eeq
Note that $\Sigma$
is a tensor product of Pauli matrices so it satisfies

\beq
\Sigma^\dagger = \Sigma, \;\; \Sigma^2=1
\;.
\label{Sigma_ids}
\eeq
Taylor expanding and using Eqs.(\ref{Sigma_ids}),
one can easily show that

\beq
Q=e^{i\frac{\pi}{4}\Sigma}
=\frac{1 + i\Sigma}{\sqrt{2}}
\;.
\eeq
Therefore\footnote{Ref.\cite{xanadu} credits
Ref.\cite{japan} as the first
paper to use
Eq.(\ref{japan_eq})
in the context of evaluating
gradients of cost functions in quantum computing.}

\beqa
B_{\tau,k} &=& i\langle H_\tau \Sigma - h.c.\rangle\\
&=&
\left\langle
Q^\dagger H_\tau Q
- Q H_\tau Q^\dagger
\right\rangle
\;.
\label{japan_eq}
\eeqa
Eq.(\ref{japan_eq}) expresses type B mean
values in a form
that is readily evaluated empirically on a real
qc. Can we do the same for type A mean values?


Recall that if $\sigma_Z(\beta)$ is the $Z$ Pauli matrix,
and $n(\beta)=\ket{0}\bra{0}_\beta$ is the number operator,
acting on qubit $\beta$,
then

\beq
1-2n(\beta) = (-1)^{n(\beta)} = \sigma_Z(\beta)
\;.
\eeq
This result relies on the fact
that the projection operator $n$
satisfies $n^2=n$  and therefore
can only have two eigenvalues, 0 and 1.
Let

\beq
\eta = n(1)n(2)
\;.
\eeq
Eta's square also equals itself so
$\eta\in \{0, 1\}$. Therefore

\beq
1-2\eta = (-1)^{\eta} = (-1)^{n(1)n(2)} = \sigma_Z(1)^{n(2)}=
\sigma_Z(2)^{n(1)}
\;.
\eeq
and

\beqa
H\eta + \eta H &=&
\frac{-1}{2}(1-2\eta)H(1-2\eta) + \frac{1}{2}H
+ 2\eta H\eta\\
&=&
\frac{-1}{2}(-1)^\eta)H(-1)^\eta + \frac{1}{2}H
+ 2\eta H \eta
\;.
\eeqa
This allows us to express type A mean values as

\beq
A_\tau =\frac{-1}{2}
\langle (-1)^\eta H_\tau (-1)^\eta\rangle
+ \frac{1}{2}\langle H_\tau \rangle
+ 2\langle\eta H_\tau \eta\rangle
\;.
\label{A-as-3-mean-values}
\eeq

Of the 3 mean values on the RHS of
Eq.(\ref{A-as-3-mean-values}),
the first two are readily
evaluated empirically on a qc.
The third one,
$\langle\eta H_\tau \eta\rangle$, is not so ready.
One possible way to evaluate
$\langle\eta H_\tau \eta\rangle$ on a qc
is to first express
the Hermitian operator $H_\tau$
as a ``QubitOperator". {\tt QubitOperator}
is a class
in the open-source software
OpenFermion. In other words, decompose
$\langle\eta H_\tau \eta\rangle$
into a linear combination
 with real coefficients $c_r$
of tensor products of a Pauli operator
(or the identity) acting on each qubit:

\beq
H_\tau = \sum_r c_r \prod_{\beta=0}^{N_b-1}
\sigma_{d_{\beta,r}}(\beta)
\;.
\eeq
Here $d_{\beta,r} \in \{0,1,2,3\}$
so as to include the identity and the 3 Pauli matrices.
It follows that

\beqa
\left\langle\eta H_\tau\eta \right\rangle &=&
\sum_r c_r
\left\langle\eta
\prod_{\beta=0}^{N_b-1}
\sigma_{d_{\beta_r}}(\beta) \eta\right\rangle
\\
&=&\sum_r c_r
\bra{1}\sigma_{d_1}(1)\ket{1}
\bra{1}\sigma_{d_2}(2)\ket{1}
\left\langle n(1)n(2)
\prod_{\beta\neq 1,2}
\sigma_{d_\beta}(\beta)\right\rangle
\;.
\eeqa
This final form for $\langle\eta H_\tau \eta\rangle$
makes it clear how to
evaluate it empirically on a qc. It involves
post-selecting the outcomes of qubits 1 and 2.

\begin{thebibliography}{99}
\bibitem{japan}
	Kosuke Mitarai, Makoto Negoro,
Masahiro Kitagawa, and Keisuke Fujii.
 ``Quantum circuit learning" arXiv:1803.00745.
\bibitem{xanadu}
PennyLane documentation has large list of references
to arXiv papers.
\end{thebibliography}

\end{document}